---
title: "fig_datasourcecompare"
author: "Leslie Roberson"
date: '2022-06-16'
output: html_document
---

```{r setup, include=FALSE}

library(here)
library(tidyverse)
library(pillar)
library(viridis)
library(scales)

knitr::opts_chunk$set(echo = TRUE)
```

#WebFigure 3
Lollipop/butterfly plot with 4 facets:
Blue shark, silky shark, tunas and billfishes, other elasmobranchs

Mean difference in recorded catch per hook between observer and logbook data for the 22 vessels occurring in both data sets. Only 15 vessels had more than 5% observer coverage. Vessels with more records on top. Positive difference (circles) means observers’ average catch rate was higher; triangles signify higher catch rates in the vessels’ logbook. 

```{r webfig3 plot}

webfig3_dat <- read_csv(here("public_data/webfig3_dat.csv"))

## prep data for plot
glimpse(webfig3_dat)

webfig3_dat <- webfig3_dat %>%
  mutate(catch_grp=factor(catch_grp, 
                          levels=c("blue_shark", "silky_shark", "tunas_and_billfishes", "other_sharks_and_rays"),
                          labels=c("Blue shark", "Silky shark", "Tunas and billfish", "Other elasmobranchs")))
webfig3_dat <- webfig3_dat %>%
  mutate(posneg_set=case_when(
    diff_avg_set>0~"obs higher",
    diff_avg_set<0~"log higher",
    diff_avg_set==0~"both zero",
    TRUE~""),
    posneg_hks=case_when(
    diff_avg_1000hks>0~"obs higher",    
    diff_avg_1000hks<0~"log higher",
    diff_avg_1000hks==0~"both zero",
    TRUE~""),
    posneg_set=factor(posneg_set, c("obs higher", "log higher", "both zero", "")),
    posneg_hks=factor(posneg_hks, c("obs higher", "log higher", "both zero", "")))
glimpse(webfig3_dat)
# make and order factors
webfig3_dat$label_nsets <- factor(webfig3_dat$count_sets)
webfig3_dat <- webfig3_dat %>%
  mutate(vesselID=factor(vesselID),
         vesselID=fct_reorder(vesselID, count_sets))
levels(webfig3_dat$vesselID)
levels(webfig3_dat$label_nsets)
webfig3_dat <- webfig3_dat %>%
  mutate(label_nsets=fct_reorder(label_nsets, count_sets)) # somehow they still don't order right in plot
           
##<>><< ##<><>>< Plot facet butterfly/Lollipop plot

webfig.3 <- webfig3_dat %>%
  # show only the 22 vessels with some observer coverage
  filter(obs_cover_pcnt>0) %>%
  ggplot() +
  geom_hline(yintercept=0, color="lightgray") +
  geom_segment(aes(x=reorder(vesselID, count_sets), xend=reorder(vesselID, count_sets), 
                   y=0, yend=diff_avg_1000hks, color=obs_cover_pcnt), size=0.5) +
  geom_point(aes(x=reorder(vesselID, count_sets), y=diff_avg_1000hks, color=obs_cover_pcnt, shape=posneg_hks)) +
  facet_wrap(~catch_grp, scales="free_x") +
  scale_x_discrete(breaks=webfig3_dat$vesselID, labels=webfig3_dat$label_nsets) +
  scale_color_viridis_c(option="plasma",
                        limits=c(0,20),
                        breaks=seq(0,20,5),
                        guide = guide_colorbar(direction = "horizontal",  title.position = "top")) +
  scale_shape(labels=c('Observer counts higher', 'Logbook counts higher'),
              guide = guide_legend(direction = "vertical", title.position = "top")) +
  coord_flip() +
  labs(y="Difference in counts per 1000 hooks", x="Recorded sets per vessel", color="Sets observed (%)", shape="") +
  theme_bw() +
  theme(
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank(),
        panel.grid=element_blank(),
        strip.background = element_rect(fill="white"),
        legend.position="top",
        legend.key.height = unit(dev.size()[1] / 55, "inches"), # reduce thickness of color bar in legend
        legend.title=element_text(size=8),
        legend.text = element_text(size=8),
        legend.margin=margin(t=-0.1,b=-0.1, r=0.5, l=0.5, unit='cm')
        )
webfig.3

ggsave(plot=webfig.3, 
       filename=here("webfig3.png"),
       dpi=300,
       device=NULL)

```

